\documentclass[12pt,oneside]{book}

\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{booktabs}
\usepackage{natbib}
\geometry{a4paper, margin=1in}

% --- Theorem Environments ---
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{remark}{Remark}[chapter]

% --- Custom Macros ---
\newcommand{\E}{\mathbb{E}}
\newcommand{\KL}{D_{\mathrm{KL}}}
\newcommand{\Filt}{\mathcal{F}}
\newcommand{\Normal}{\mathcal{N}}
\newcommand{\Gam}{\text{Gamma}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\R}{\mathbb{R}}

\title{Bayesian Approaches to Sequential Decision-Making \\ in Uncertain Environments}
\author{Ansh Dani \\ Barrett, The Honors College \\ Arizona State University}
\date{May 2026}

\begin{document}

\maketitle

% ---------------------------------------------------------------
%  ABSTRACT
% ---------------------------------------------------------------
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

Sequential decision-making under uncertainty requires agents to simultaneously
learn unknown environment parameters and exploit current beliefs to maximize
long-run wealth—the classic exploration-exploitation dilemma. Classical
solutions, such as the Kelly Criterion and naive Bayesian plug-in estimators,
are asymptotically optimal in stationary environments but fail catastrophically
when the underlying return distribution shifts without warning, a phenomenon we
term the \emph{Sticky Prior Paradox}: posterior mass accumulated under one
regime resists rapid reallocation when the true regime changes, causing agents
to over-bet into an adversarial market.

This thesis makes three interrelated contributions.  First, we establish a
rigorous simulation framework—grounded in measure-theoretic probability and the
Strong Law of Large Numbers—that enforces strict filtration constraints and
Common Random Numbers methodology to enable statistically valid agent
comparisons.  Second, we propose a \emph{Volatility-Augmented Hidden Markov
Model} (Vol-HMM) that augments the standard univariate return observation with
a rolling volatility feature and a CUSUM change-point trigger, reducing regime
detection lag from approximately 15--20 time steps under sticky-prior agents to
under 2 steps.  Third, we implement a \emph{Risk-Constrained Kelly} strategy
via Constant Proportion Portfolio Insurance (CPPI), which provably bounds the
maximum drawdown at a user-specified floor while quantifying the exact ``cost
of survival'' in foregone geometric growth.

We validate all three contributions across three synthetic environments
(stationary, slow-drift, adversarial shock) and four empirical S\&P~500 epochs
(2000 Dot-Com Collapse, 2008 Global Financial Crisis, 2020 COVID Crash, 2013--19
Bull Market), incorporating realistic transaction costs.  Fisher's Exact Tests
confirm that the Vol-HMM achieves statistically significant reductions in ruin
probability relative to adversarial and Bayesian baselines
($p < 0.05$) under the harshest shock regimes.  Sharpe and Sortino ratio
analyses further demonstrate that the CPPI floor improves risk-adjusted returns
during crisis epochs without sacrificing median wealth in stable environments.

\bigskip
\noindent\textbf{Keywords:} Kelly Criterion, Multi-Armed Bandits, Thompson
Sampling, Hidden Markov Model, CPPI, Regime Detection, Bayesian Sequential
Decision-Making, Non-Stationary Environments.

\tableofcontents

\include{00_notation}
\include{01_introduction}
\include{02_literature}
\include{03_foundations}
\include{04_methodology}
\include{05_experiments}
\include{06_conclusion}

\include{appendix_A_code}

\end{document}
